---
title: "Item demand forecast challenge"
author: Yuri Matheus Dias
date: "`r format(Sys.time(), '%d/%m/%Y')`"
---

# Item demand competition

My goal with this notebook is to give a general EDA on the data, and maybe execute a sample model or two for prediction.
I want to learn how time series works, and train my R knowledge with this.

# Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table) 
library(lubridate)
library(parallel)
library(gridExtra)
```

Let's import the data!

```{r}
data.train = fread('../input/train.csv', header = TRUE, sep = ',')
data.test = fread('../input/test.csv', header = TRUE, sep = ',')
submission.sample = fread('../input/sample_submission.csv', header = TRUE, sep = ',')
```

Let's work only on the training data and ignore the pre-selected test data for now.

```{r}
summary(data.train)
```

We can transform that data into lubridate's format for easier handling.
We can also make `store` and `item` factors, as we know that they are just identificators from the overview page.

```{r}
#data.train$date = lubridate::ymd(data.train$date)
data.train[,
   `:=`(
     date = ymd(date),
     store = as.factor(store),
     item = as.factor(item)
   )
]
summary(data.train)
str(data.train)
```

Great, this also shows us that there are no missing values and nothing else to transform for now.
Let's take a look into the data.

I'll take a note from other kernels and add these features:
```{r}
data.train[,
   `:=`(
     weekday = lubridate::wday(date, label = TRUE, abbr = TRUE),
     month = month(date, label = TRUE, abbr = TRUE),
     year = year(date)
   )
]
```


# EDA

First, total sales per day.

```{r}
total.sales = data.train %>%
  group_by(date) %>%
  summarise(total = sum(sales))

ggplot(
  data = total.sales,
  aes(x = date, y = total)
) +
  geom_line()
```

There is definitely seasonality and trend in this.

```{r}
total.sales.store = data.train %>%
  group_by(date, store) %>%
  summarise(total = sum(sales))

ggplot(
  data = total.sales.store,
  aes(x = date, y = total, color = store)
) +
  geom_line()
```

Pretty, but doesn't tell us much, other than some stores sell more than others.

Let's try another way:

```{r}
ggplot(
  data = data.train %>%
    group_by(store) %>%
    summarise(total = sum(sales)),
  aes(x = store, y = total, fill = store)
) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
    labels = scales::format_format(
      big.mark = ".", decimal.mark = ",", scientific = FALSE
    ) # it's absolutely ridiculous how this isn't already built-in!
  )
```

Some stores generate almost double the amount of sales in total!
#7 there is not looking so good.

Let's do the same for products.

```{r}
ggplot(
  data = data.train %>%
    group_by(item) %>%
    summarise(total = sum(sales)),
  aes(x = item, y = total, fill = item)
) +
  geom_bar(stat = 'identity') +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1.0, vjust = 1.0)
  ) + 
  guides(fill = FALSE, color = FALSE)
```

Chaotic, but hey! Colors! Some items sell almost 3x as much as others, so that's interesting to see with the plots too.

